Enabling Ambient Backscatter \\Using a Low-Cost Software Defined Radio

### 2017-05-15 abstract.tex
Samll changes in abstract and fig. 5.

Ambient backscatter communication enables ultra-low
power wireless transmissions and is hence attractive
for networking devices which operate on harvested energy. 
The data to be sent is modulated on an ambient signal, 
such as TV signals. Tag-to-tag backscatter communication 
requires rather strong ambient signals to achieve acceptable
error rates and range.
In this paper, we present a design and implementation of a
backscatter receiver on a software defined radio plattform 
(RTL-SDR), achieving higher sensitivity than state-of-the-art
backscatter tags and thus enabling wide area ambient backscatter 
communication from low-power tags. 
Our measurements show that the TV signal is strong 
enough in most parts of a mid-sized Swedish city for our
system to operate and communicate.

>>>>

Ambient backscatter communication enables ultra-low
power wireless transmissions and is hence attractive
for networking devices which operate on harvested energy. 
The data to be sent is modulated on an ambient signal, 
such as TV signals. Tag-to-tag backscatter communication 
requires rather strong ambient signals to achieve acceptable
error rates and range.
In this paper, we present a design and implementation of a
backscatter receiver on a extremely low-cost software defined radio plattform 
(RTL-SDR), achieving higher sensitivity than state-of-the-art
backscatter tags and thus enabling wide area ambient backscatter 
communication from low-power tags. 
Our measurements show that the TV signal is strong 
enough in most parts of a mid-sized Swedish city for our
system to operate and communicate.

### 2017-05-15 paper.tex
Samll changes in abstract and fig. 5.

Backscatter operates by reflecting or absorbing impinging
radio signals on the antenna. To purposefully toggle the antenna
to achieve these states, we  terminate the antenna to a 
RF switch. The RF switch enables us to alternate the impedance connected to
the antenna between matched (50 \ensuremath{\Omega}) and open state.
We control the RF switch through an I/O pin on a low-powered MCU, Texas
instruments MSP430. As discussed earlier, to reduce the effects of
self-interference, we frequency-shift the backscatter transmission
away from the TV signal. Hence, we operate the RF switch using a 
\SI{2}{\mega\hertz} intermediate frequency signal. Therefore the result 
is a frequency shift of the television signal by 2 MHz, when
transmitting a bit 1, and no shift for a 0.  Hence,  we 
leverage amplitude modulation at receiver in our design.
binary amplitude shift keying is the implemented modulation technique.

@@

The resulting values are sent through the bus to the decoder. After the decoder
received the samples it decides when a frame is sent or when the channel is
idle. The decoder also includes a function to correlate the received bitstream
with an expected pattern. Hence it is able to determine the bit error ratio
(BER). Furthermore we have implemented a series of aditional tools that include
different simulators for e.g. playing back recorded data and a tool (an
oscilloscope) to plot the data, which is written into files by the demodulator.
An example of a recorded data stream with 25 kS/s can be seen in Figure
\ref{fig:transmission}.

@@

\includegraphics[width=\columnwidth]{./fig/transmission}
\vspace{-30mm}
\caption{Start of a transmission of 101010.. with 1 kbit/s with idle state before. Data from the Octave oscilloscope. Amplitude is quite low with an average of 12 \% of the maximum, which is 127.5. This is due to a low signal strength. Sampling frequency is 25 kHz.}
\label{fig:transmission}
\vspace{-6mm}

>>>>

Backscatter operates by reflecting or absorbing impinging
radio signals on the antenna. To purposefully toggle the antenna
to achieve these states, we  terminate the antenna to a 
RF switch. The RF switch enables us to alternate the impedance connected to
the antenna between matched (50 \ensuremath{\Omega}) and open state.
We control the RF switch through an I/O pin on a low-powered MCU, Texas
instruments MSP430. As discussed earlier, to reduce the effects of
self-interference, we frequency-shift the backscatter transmission
away from the TV signal. Hence, we operate the RF switch using a 
\SI{2}{\mega\hertz} intermediate frequency signal. Therefore the result 
is a frequency shift of the television signal by 2 MHz, when
transmitting a bit 1, and no shift for a 0.  Hence,  we 
leverage amplitude modulation at receiver in our design.

@@

The resulting values are sent through the bus to the decoder. After the decoder
received the samples it decides if a frame is received or if the channel is
idle. The decoder also includes a function to correlate the received bitstream
with an expected pattern. Hence it is able to determine the bit error ratio
(BER). Furthermore we have implemented a series of aditional tools that include
different simulators for e.g. playing back recorded data and a tool (an
oscilloscope) to plot the data, which is written into files by the demodulator.
An example of a recorded data stream with 25 kS/s can be seen in Figure
\ref{fig:transmission}.

@@

\includegraphics[width=\columnwidth]{./fig/transmission}
\vspace{-30mm}
\caption{Data from the Octave oscilloscope showing the start of a transmission of 101010.. with 1 kbit/s with idle state before. The amplitude is quite low with an average of 12 \% of the maximum, which is 127.5. This is due to a low signal strength. Sampling frequency is 25 kHz. At the top one can see the raw signal from the ADC, in the middle the filtered signal and at the bottom the signal after the Schmitt trigger.}
\label{fig:transmission}
\vspace{-6mm}
